---
title: "State of the API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{State of the API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

There's a lot going on with the new version of the Patentsview API, recently renamed
PatentSearch API, as announced [here](https://search.patentsview.org/docs/#naming-update).

## On the Plus Side

* All fields can be queried now.

* The 100,000 row result set size limit seems to be gone.  Might be part curse as a user might not have
intended to [retrieve all utility patents](api-changes.html#unbounded) etc.

* The new version of the API returns data in types other than strings, e.g. application.rule_47_flag from the patent endpoint is a boolean.  ```View(fieldsdf[grep('number|integer|boolean',fieldsdf$data_type),c("endpoint","field","data_type")])``` currently returns 90 rows

  There are still a few fields where the API returns strings that cast-pv-data() casts to a more appropriate data type[^7].
fieldsdf$data_type is set to int or bool for these fields.

  ```{r}
library(patentsview)

# Count of non string, non date fields returned by the new version of the API
nrow(fieldsdf[grep("number|integer|boolean", fieldsdf$data_type), ])

print(fieldsdf[grep("bool$|int$", fieldsdf$data_type), c("endpoint", "field", "data_type")], row.names = FALSE)
  ```
* The API team created a [Swagger UI](patentsview-breaking-change.html#online-documentation) page for the new version of the API. On the [Joys of Swagger UI](https://patentsview.historicip.com/swagger/articles/)

## Issues

* The original API's sunset date is currently February 12, 2025 though the new version 
still has [issues](api-changes.html#things-to-note)
* On a new [string case sensitivity](api-changes.html#case-sensitivity-caveat), opened as a [bug](#case-dependent) listed below
* On [string vs full text operators](api-changes.html#string-and-full-text-operators), also opened as a bug listed below.
Note that there isn't a way to distinguish strings and full text fields in the OpenAPI object
(limitation of the OpenAPI spec).  We may need to change the way fieldsdf is generated
if we need to track what operators can be used, possibly parsing the [API's endpoints page](https://search.patentsview.org/docs/docs/Search%20API/SearchAPIReference/#endpoints)

* API Oddities
   + Some [requested fields](api-changes.html#slight-weirdness) ending with _id come back without the _id[^4] 
Ex patent endpoint requested inventors.inventor_id and
assignees.assignee_id are returned as inventors.inventor and assignees.assignee, their values are
their respective HATEOAS links

   + Two of the [HATEOAS links](api-changes.html#hateoas-links) use a colon instead of having two URL parameters
      * https[]()://search.patentsview.org/api/v1/cpc_group/G01S7:4865/
      * https[]()://search.patentsview.org/api/v1/uspc_subclass/403:57/ (endpoint currently throws a 500)

      An [API key](api-changes.html#an-api-key-is-required) is now required so, intentionally, these URLs aren't clickable since no API key would be sent, resulting in a 403, Forbidden response.

   + Field inconsistencies
      * There are two rule_47_flag fields, one returned by the patent endpoint and one returned by the
publication endpoint.  The former is returned as a boolean, the latter as a string that cast-pv-data casts to a boolean.[^1]

      * assignee_type, available from three endpoints, is returned as a string but looks like they're integer values[^3]

      * Most document_number fields are integers but there are two that are returned as strings
that cast-pv-data casts to integers.[^2] [Additional details](getting-started.html#casting-fields)

   + Most endpoints' returns are the plural form of the singular endpoint with these exceptions: 

     |endpoint|returned entity|
     |--------|---------------|
     |ipc|ipcr|
     |otherreference|other_references[^5]|
     |wipo|wipo|
     |publication/rel_app_text|rel_app_text_publications| 
     (Note that patent/rel_app_text returns rel_app_texts, following the singular/plural pattern)

   + Not exactly an oddity, but with no better place for this, here's a comparison of the old and 
new attributes that can be sent to the API in its o: (options) parameter via search_pv()

     |Original API|New Version| Purpose|
     |------------|-----------|--------|
     | per_page   | size      |  maximum number of rows to return |
     | page       | after     |  page through large result sets |
     | subent_cnts|           |  whether the query results should include the total counts of unique subentities|
     | mtchd_subent_only|     | whether a query should return all related subentities or just those that match query criteria.|

## Open API Bugs

Weirdly, you can only view bugs you've submitted.  I'm assuming there are other open bugs.

PVS-1147 <a name="case-dependent">	
Results are case dependent now when using an implied or explicit equals

PVS-1125	
Not all the fields in the OpenAPI object can be requested

PVS-1109 <a name="otherreference">  
The otherreference endpoint rejects the default [Swagger UI](https://search.patentsview.org/swagger-ui/) parameters (throws a 400 Bad Request Error if
either reference_sequence or reference_text is requested) and returns no data when only
patent_id is requested.  The OpenAPI object says the returned object is other_references which 
is another exception to the singular endpoint/plural return pattern[^6].

PVS-1155  
Documentation inconsistencies  
The endpoint is listed as /api/v1/attorney/, it should be /api/v1/patent/attorney for the GET/POST and /api/v1/patent/attorney/{attorney_id}/ for the GET with a url parameter
The beta endpoints say they are only GETs. The Swagger UI page and OpenAPI object say they accept posts too, which do work.

PVS-1218	
openapi.json errors
  
* "publication","us_parties.applicant_authority" has a type of "keyword" when it should be "string"

* There are two rule_47_flag fields, one is a boolean (line 7019), the other is a string (line 8004) 

* Most of the document_numbers are integers but from publication/rel_app_text it's a string as is citation_document_number from patent/us_application_citation

PVS-1306
The API accepts invalid fields  
The API accepts invalid fields that start out looking like valid fields when it should throw an error. Ex f: is["patent_iddddddddddddd", "patent_dateagogo"] and q: is {"patent_idd":"10000000"} with this result: { "error": false, "count": 0, "total_hits": 0, "patents": [] }

PVS-1181
Improvement Suggestion  
There isn't a data dictionary for the API like there is for the bulk download files.
A specific question would be what is the difference between patent_earliest_application_date and application.filing_date returned by the new patent endpoint. 
Other questions would be what do the values of the assignees.assignee_type field represent, are they all integers and if so should the field be received as an integer rather than a string?

## State of the R Package

* [Two new methods](api-changes.html#additional-r-package-changes) were added to the R package

* The reworked [ropensci post](ropensci_blog_post.html) explains what changes had to be made since assignee 
latitude and longitude are no longer available from the patent endpoint.

* A new vignette, inspired by the API team's jupyter notebook, was created showing two of the things it mentions

  1. The API and now the R package allow nested attributes to be [wildcarded](understanding-the-api.html#fields-shorthand),
where the group's name will retrieve all the group's nested fields.  An empty string can also be passed as a group name to get_fields()
which then requests all the non-nested attributes from the endpoint. Ex. ```get_fields("patent", group=c("", "inventors")```

  2. On [applying multiple conditions to related-entity fields](understanding-the-api.html#unexpected-results)

* The endpoints are now singular, to match the API's change.  Some of the new endpoints are nested under patent/ and one is nested under publication/,
also matching the API's change
  ```{r}
print(unique(fieldsdf[grep("/", fieldsdf$endpoint), "endpoint"]), row.names = FALSE)
  ```
* Under the covers, httr was changed to httr2 (see several mentions below).  This only matters to users if 
they passed additional arguments (...) to search_pv().  Previously if they passed config = httr::timeout(40)
they'd now pass timeout = 40 (name-value pairs of valid curl options, as found in curl::curl_options())

* Now that the R package is using httr2, users can make use of its last_request() method to see what was sent to the API.  This could be useful when trying to fix an invalid request.  Also fun would be seeing the raw API response.
```
httr2::last_request()
httr2::last_response()
httr2::last_response() |> httr2::resp_body_json() 
```

* Package contributor specific changes

    + Changed httr to httr2 as [httr](https://httr.r-lib.org/index.html) says it's superseded by [httr2](https://httr2.r-lib.org/)
It's trying to put us out of business by handling 429 retries etc.  See their 
[Wrapping APIs vignette](https://httr2.r-lib.org/articles/wrapping-apis.html)

  + My github repo is set to deliver to https://mustberuss.r-universe.dev/patentsview.  The new
version of the R package can be installed via   
```install.packages('patentsview', repos = c('https://mustberuss.r-universe.dev'))```

  + On the new [implementation of paging](api-changes.html#a-note-on-paging) and the
[PR discussion](https://github.com/ropensci/patentsview/pull/29#discussion_r1059153136) on paging with more than a primary sort.
This added a dependency to data.table.

  + The patent/otherreference endpoint isn't currently working (reported as a [bug](#otherreference) above).
It is included in the return of get_endpoints() and has only a negative test case that will
fail when the API responds with something other than an error.

  + The R package unit tests have some skips and other negative tests that API bugs still exist
(they'll fail when bugs are fixed).  The tests were copied to
[a single file](https://github.com/mustberuss/patentsview/blob/api-redesign/tests/testthat/test-api-bugs.R) that
was then submitted to the API team as bugs.  Most of the tests exist
in other test files but a few are unique (specific to API bugs that won't be needed
once they're fixed).

  + A possible [tech note](patentsview-breaking-change.html) for when the new package is ready mentions

    > As shown in the updated [Top Assignees vignette](top-assignees.html), there will be occasions now where multiple API calls are needed to retrieve the same data as in a single API call in the original version of the API and R package.

  + The workflow (R-CMD-check.yaml) was updated to get rid of deprecated warnings

  + Added vignettes/build_some.R to half render one or more of the vignettes during development

  + Added a ```bootstrap: 5``` to _pkgdown.yml so the search box would display (Maëlle's [PR](https://github.com/ropensci/patentsview/pull/34))

  + The R package handles [throttling](api-changes.html#throttling), which the new version of the API now imposes

  + data-raw/fieldsdf.R is working again!  There is some hard-coding that can be removed when
bugs are fixed.  We may need to retain some of the hard-coding, see the comments in data-raw/fieldsdf.R

  + <a name="improvements"> Possible Package Improvements

    * The version number ought to be bumped to 1.0.0 since there are breaking API changes (singular endpoints and the addition of an API key).  validate-args.R has some version
specific code that may need modifying.
    * Result set size seems unbounded now.  Should we warn if a query would return more than 100,000 rows with all_pages = TRUE?
    * Have get_fields() and search_pv() throw a specialized error if a plural endpoint is passed
    * Add an issue template that warns users not to share their API key
    * Add a contributing.md or something that explains how to build everything, something like
[Findings/Contributor 101](#findingscontributor-101) below
    * Navigation on the vignettes could be better,  [understanding-the-api](understanding-the-api.html) isn't a link in the navigation yet,
neither is the possible [tech note](patentsview-breaking-change.html)
    * API attribute changes page and per_page to be explained somewhere.
    * httr2 improvements
       + Can the throttled test detect output to stdout etc?  We used to expect_message "The API's requests per minute limit has been reached." Now "Waiting 45s for retry backoff" appears but doesn't satisfy expect_message().  Currently using system.time()


## Worth Monitoring

bergant/rapiclient has a new maintainer and [he's in favor](https://github.com/bergant/rapiclient/issues/11#issuecomment-2383483280)
of supporting OpenAPI v3 though he thinks that would warrant a new package.   The R version of the fieldsdf creator
is working again!  It uses rapiclient which is expecting to read a Swagger/OpenAPI v2 object
(the patentsview object is OpenAPI v3).  It throws a warning but seems to work otherwise.

## Findings/Contributor 101

Helpful info after you've forked https://github.com/ropensci/patentsview 

* API related

  + Updated versions of the new API are unannounced for the most part, maybe on a three week release cycle, often on 
Saturdays.  I do head requests from time to time on the [API's OpenAPI object](https://search.patentsview.org/static/openapi.json)
to detect when its Last-Modified changes. When it changes, I back up data-raw/fieldsdf.csv, run data-raw/fieldsdf.R and 
compare the backed up fieldsdf.csv to the new one to see if anything changed. If it did change, I'll do a
local build and run all of the tests to see what happens and push any changes to keep the tests passing.
Quarterly or so they'll update https://patentsview.org/release-notes

  + Report API suggestions/bugs etc [here](https://patentsview-support.atlassian.net/servicedesk/customer/portals)
or Community -> Support using the nav on [patentsview.org](httsp://patentsview.org)

  + Now all the endpoints are documented on a [single page](https://search.patentsview.org/docs/docs/Search%20API/SearchAPIReference/#endpoints).
The [query language]( https://search.patentsview.org/docs/docs/Search%20API/SearchAPIReference/#api-query-language) is also on that same page.
Originally there was a separate page for each endpoint.

  + The patentsview forum isn't terribly active but it's worth keeping an eye on
https://patentsview.org/forum

* Locally

  + Build and install the local code ```devtools::install('.')```

  + rscript data-raw/fieldsdf.R to run the fieldsdf creator locally.  It parses the Patentsview OpenAPI object to 
produce fieldsdf.csv and fieldsdf.rda

  + Local Testing

    * devtools::test()  to run all the tests in tests/testthat/

    * devtools::test(filter="utils") to just run just test-utils.R

    * note that if you add new test methods, they must start with ```skip_on_cran()``` or builds will fail on r-universe (see Optional below) and eventually on CRAN.

  + Half building vignettes locally - see vignettes/README.md

    * need phantomjs in your path on windows at least

    * see the comments in vignettes/build_some.R 

  + pkgdown locally

    * pkgdown::build_articles() to see local changes to half-rendered vignettes 

    *  You may want to run individual methods or 
[pkgdown::build_site()](https://pkgdown.r-lib.org/reference/build_site.html) which is a wrapper for 
       + init_site()
       + build_home()
       + build_reference()
       + build_articles()
       + build_tutorials()
       + build_news()
       + build_redirects()
<br />
<br />
   + build reference pages locally   
If you make changes to method documentation, run 
      * devtools::document() and
      * pkgdown::build_reference()
<br />
<br />
   + see README.Rmd changes locally
      * knitr::knit("README.Rmd", "README.md")
      * pkgdown::build_home()
<br />
<br />
* Remotely

  + pkgdown remotely   
   You can set up github pages in your repo to display vignettes and man pages, e.g. https://github.com/mustberuss/patentsview

  + Remote Testing   
You can put your API key in your repo as a Repository secret so the tests will run when you push changes.
 E.g., I have PATENTSVIEW_API_KEY set in 
https://github.com/mustberuss/patentsview/settings/secrets/actions (though settings aren't publicly
viewable)

* Before doing a push 

  + make sure you are following the [Tidyverse style guide](https://style.tidyverse.org/).
There are plugins to keep you stylish ```styler::style_pkg(".", dry = "fail")```
```styler::style_file('path/file')``` if necessary or use a plugin etc

   + make sure the examples still run ```devtools::run_examples(run_dontrun = TRUE)```

   + check spelling using ```devtools::spell_check(pkg = ".", vignettes = TRUE, use_wordlist = TRUE)```
Locally I have an unchecked in inst/WORDLIST with additional words to allow (one word per line)
see https://devtools.r-lib.org/reference/spell_check.html

* Optional

  + you can set up your own r-universe by creating a special github repo, as explained 
[here](https://ropensci.org/blog/2021/06/22/setup-runiverse/).
E.g., https://github.com/mustberuss/mustberuss.r-universe.dev
with https://github.com/mustberuss/mustberuss.r-universe.dev/blob/master/packages.json
See [this article](https://ropensci.org/blog/2023/02/07/runiverse-registry-repo/) on the newer repo naming convention
<br />
<br />
* Notes
   + By default, patentsview is on r-universe https://ropensci.r-universe.dev/patentsview
   + Someone really smart should write a python library for the API!
<br />
<br />  
    
* Questions:
   + Is there a way to set up notifications on r-universe build failures?  If you miss a
skip_on_cran(), tests could work in your repo (if you have your API key as a repo secret)
but the build will silently fail on r-universe.  Don't ask how I know that.

   + Why don't the vignettes show up on r-universe?  The articles link is grayed out.

      Ah, on the build it shows Build pkgdown site (ropensci only) was skipped

   + Should we add a row_limit or something?  We'd page our way and stop when the row_limit is met.  For someone wanting more than a 1000 rows but not necessarily all the rows, especially since
there isn't the 100,000 row limitation now.  The API's ```after``` is now exposed 
in search_pv() so users could do their own paging.

[^1]: Observation sent to the API team.
[^2]: Observation sent to the API team.
[^3]: There isn't a data dictionary for the API so I suggested they create one.
[^4]: Observation sent to the API team but they might not see it as a bug.
[^5]: To be fixed by the API team
[^6]: To be fixed by the API team
[^7]: Observation sent to the API team.
